Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-10-04 paper/paper.tex
added another paragraph to Discussion

This viral genetic information turns out to be extremely important for understanding infection outcome and innate-immune induction.
Despite the fact that we used a low-passage viral stock generated from plasmids encoding a ``wildtype'' influenza genome, a large majority of cells are infected by virions that have mutations or fail to express a gene.
Although our study is certainly not the first to note that influenza virus has a high mutation rate~\citep{parvin1986measurement, suarez1992heterogeneity, suarez1994estimation, bloom2014experimentally, pauly2017novel} and sometimes fails to express genes~\citep{brooke2013most, heldt2015single, dou2017analysis, russell2018extreme}, it is the first to directly observe the full spectrum of these defects across single cells.
Simple visual inspection of \FIG{genotypes} makes it glaringly obvious how any experimental approach that does not directly sequence the viral genes in single cells is really averaging across a vast spectrum of hidden viral defects.

>>>>

This viral genetic information turns out to be extremely important for understanding infection outcome and innate-immune induction.
Despite the fact that we used a low-passage viral stock generated from plasmids encoding a ``wildtype'' influenza genome, a large majority of cells are infected by virions that have mutations or fail to express a gene.
Although our study is certainly not the first to note that influenza virus has a high mutation rate~\citep{parvin1986measurement, suarez1992heterogeneity, suarez1994estimation, bloom2014experimentally, pauly2017novel} and sometimes fails to express genes~\citep{brooke2013most, heldt2015single, dou2017analysis, russell2018extreme}, it is the first to directly observe the full spectrum of these defects across single cells.
Simple visual inspection of \FIG{genotypes} makes it glaringly obvious how any experimental approach that does not directly sequence the viral genes in single cells is really averaging across a vast number of hidden viral defects.

It turns out that these idiosyncratic viral defects explain much of the heterogeneity in viral transcriptional burden that has been observed in prior single-cell studies of influenza~\citep{russell2018extreme, steuerman2018dissection, heldt2015single, sjaastad2018distinct}.
If we examine only cells infected by truly wildtype virions, we see that the heterogeneity in viral transcriptional burden is substantially reduced (\FIG{mutations}A).

Crucially, viral defects also play an important role in determining if infected cells trigger an innate-immune response.
We identified four types of viral defects in single IFN+ cells that we validated to increase IFN induction. 
Two types of defects---absence of the NS gene and amino-acid mutation to the NS1 protein---likely act by impairing NS1's well-known ability to antagonize innate-immune pathways~\citep{garcia1998influenza, hale2008multifunctional}.
The third type of defect, amino-acid mutations in the PB1 protein, is consistent with recent work by \citet{velthuis2018mini} showing that mutations to the viral polymerase affect generation of mini-viral RNAs that activate the innate-immune sensor RIG-I.
Finally, we found an internal deletion in the PB1 gene that enhances IFN induction, consistent with a large body of prior work showing that such deletions are immunostimulatory~\citep{baum2010preference, tapia2013defective, boergeling2015evidence, dimmock2015cloned}.
In fact, given the extensive prior work describing such deletions, it is perhaps surprising that we did not identify more of them in our single IFN+ cells.
There may be several reasons: first, we used relatively pure viral stocks~\citep{xue2016propagation} at modest MOI~\citep[conditions that better reflect the early stages of real human influenza infections;][]{mccrone2018stochastic, xue2018reconciling, varble2014influenza}; second, our experiments preferentially captured cells with higher viral transcriptional load; and third, most prior studies have used techniques that can detect large deletions but not subtle point mutations.
But we emphasize that the relative importance of different types of viral defects for IFN induction likely varies across infection conditions, viral strains, and cell types---so it certainly remains an open question as to which defects are most relevant for early immune detection of actual viral infections in humans.
